---
title: "Statistical Analysis Carmo et al - Horse behaviour"
author: "Ruan Daros"
date: "June/2022"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
editor_options:
  chunk_output_type: console
always_allow_html: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

## Background

Animal emotion research to date has focused more on negative than positive emotions.
But with the explicit incorporation of positive welfare into the animal welfare 
concept (Mellor, 2016) this area of research has seen significant development in 
the past decade (Boissy et al 2007; Rault et al 2019). Animal behaviour studies 
can indicate how an animal feels under a given circumstance (Weary et al., 2017) 
and have been applied to study positive emotions successfully (e.g., van der Harst 
and Spruijt, 2007; Yeates and Main, 2008).

## Objectives and hypothesis

The aims of this study are to assess the impact of positive reinforcement on horse
anticipatory behaviour to enter a horse chute and evaluate horse facial movements 
before and after receiving a food reward. Horses are expected to show marked
anticipatory behaviour (e.g., more time close to the gate that leads to the horse 
chute) during the positive reinforced phase compared to the habituation phase (no
reinforcement applied) and facial expressions should be distinct before and after
receiving a food reward.

## Methods

Data was collected in the last three weeks of February 2021. Week 1 being the
baseline and habituation phase and weeks 2 and 3 the experimental phase. Data 
was collected on the afternoons of Wed, Thu and Fri of each week. Horse behaviour
was video recorded and further tabulated. Specifically, in the experimental phase
half of the animals received positive reinforcement in week 2. The other half of 
the animals were used as control and no reinforcement was applied. In week 3 a
cross-over between the groups was applied (i.e. the half of the animals that
received the food rewards in week 2 now became the control group).

Video clips of the third day from the positive reinforcement (PR) and control
phases were analyzed to extract the data regarding the facial action units (i.e.
facial expression) of each animal. Interobserver reliability was calculated with
data collected independently by LGC and LCW on 12 60-sec video clips of horses
participating in the study. In our analysis we considered 35 action units described
in Wathan et al., (2015). Most action units were considered events but some were
considered states and duration of each one of these were calculated.

## Statiscal approach

Experimental design features include repeated measures, a baseline period and an
experimental period with cross-over. This allows for two approaches - Btw / within
subject. When repeated measures are used horse will be included as random effect
in the GLMs. There are several outcomes of interest in this study, with several 
types: binary, continuous (normal) and counts. When possible simpler statistical
tests will be applied (eg. paired t-test for facial expression data).

### Packages and custom functions
```{r, message=F, warning=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggsignif)
library(irr)
library(emmeans)
ORand95CI <- function(model, confint_method) {
  
  model_OR <- fixef(model)
  
  if (confint_method == "profile")
    model_CI <- confint(model)
  if (confint_method == "wald")
    model_CI <- confint(model, method = "Wald")
  
  table <- round(data.frame(exp(model_OR), exp(model_CI[-1, ])), digits = 2)
  names(table) <- c("OR", "LowerCI", "UpperCI")
  
  return(print(table))
}

```

### Loading dataset & Wrangling & Descriptive stats

```{r, warning=FALSE}
#loading dataset and creating variables related to time closer to the gate
raw <- read_csv("Dados_Laize_ISAE2021.csv") %>% 
  mutate(TT_BEF = AREA1_ANTES + OUTRAS_ANTES,
         TT_AFT = AREA1_DEPOIS + OUTRAS_DEPOIS,
         PropT_A1_BEF = AREA1_ANTES/TT_BEF,
         PropT_A1_AFT = AREA1_DEPOIS/TT_AFT,
         rest_alltime = case_when(DESCANSO == "30-60" ~ "1",
                                    is.na(DESCANSO) ~ "NA",
                                    TRUE ~ "0"),
         rest_never = case_when(DESCANSO == "0" ~ "1",
                                  is.na(DESCANSO) ~ "NA",
                                  TRUE ~ "0"),
         rest_part_time = case_when(rest_alltime == "1" ~ "0",
                                 rest_never == "1" ~ "0",
                                 TRUE ~ "1"))

# first day(s) of treatments has no impact on animal beh. thus we can create 
# datasets that only the last day of the experiment. we also can compare PR vs
# baseline when needed.

#creating a dataset for comparing PR versus control (only last day of each)
data_btw <- raw %>% 
  filter(EXP_DAY != 4,
         EXP_DAY != 5,
         EXP_DAY != 7,
         EXP_DAY != 8) %>% 
  filter(GRUPO != "Baseline")

# creating a dataset for baseline vs PR (only last day)
data_wth <- raw %>% 
  filter(EXP_DAY != 4,
         EXP_DAY != 5,
         EXP_DAY != 7,
         EXP_DAY != 8) %>% 
  filter(GRUPO != "Controle")

# Importing EquiFACS data
FACS <- read_csv("data_EquiFACS.csv")


# We also further collected data on freq of all FACS and recalculated frequencies # We will import this dataset as well.
FACS2 <- read_csv("data_EquiFACS_Extras_and_all_F.csv")

#We have another set with all extras FACS and the AUH13 that wasnot complete in
#the other sets
FACS3 <- read_csv("EquiFACS_AUH13_And_Extras.csv")
```

## Results

### Anticipatory behaviours

Let's check if treatment had an effect on order arrival. For this comparison we
can use all days from baseline against final two days of the PR phase.

```{r}
tmp_data <- raw %>% 
  filter(GRUPO != "Controle") %>% 
  filter(EXP_DAY != 4)

arrival_order <- lmer(ORDEM ~ GRUPO + (1|NOME), data = tmp_data)
summary(arrival_order, corr = F) #no effect of treatment on order of arrival)

rm(tmp_data)
```
**No effect of PR on order arrival**

Another behavior of interest is the number of times the animals entered the
central area of the holding area and time in the central section of the holding
area. Hypotheses: Cows anticipating a reward will enter (or spend more time in) 
the center more often. In this case we can check the last day of control and PR
phase, if they do note differ we can check experimental (PR only) vs baseline.

```{r}
# Control vs PR

#model
N_area1a <- lmer(N_AREA1 ~ GRUPO + (1|NOME), data = data_btw)
summary(N_area1a, corr = F) #no effect

# second approach experimental phase vs baseline

N_area1b <- lmer(N_AREA1 ~ GRUPO + (1|NOME), data = data_wth)
summary(N_area1b, corr = F) #still, no effect

# Time in Area 1 (proportion of time will be used to avoid confounding)

# model between exptal phase Area 1 before passing through the chute. Hypotheses:
# Horses will stay for longer in A1 before
prop_a1_befa <- lmer(PropT_A1_BEF ~ GRUPO + (1|NOME), data = data_btw)
summary(prop_a1_befa, corr = F)

#Proportion of time in area 1 before (comparison between baseline and PR)

prop_a1_befb <- lmer(PropT_A1_BEF ~ GRUPO + (1|NOME), data = data_wth)
summary(prop_a1_befb, corr = F)

```
**Most of the variables did not support the hypothesis.** However, when comparing
the last day of the PR phase against baseline (all three days), we see a tendency.
Horses spent 40% of the time in the central area before passing through the chute
while they only spent 28% of the time in the central area before passing through
the chute.

### Behaviours in the chute

Several behaviors were assessed upon horse entered the chute. Most of them were
assessed as if they happened or not, thus they are binary. We will used mixed 
log reg for those. A similar approach as used above of comparing between groups
(control/PR) and comparing between baseline and PR will be applied.

**NECK POSITION**
Lowered neck represents horses being calm and relaxed, hyper elevated neck 
represents attentive behavior. However hyper elevated neck was only seen 5
times. Thus we will not check this behavior.

```{r}
#PR vs control
lowered_necka <- glmer(NECK_ABAIXO ~ GRUPO + (1|NOME), 
                    data = data_btw,
                    family = binomial)
summary(lowered_necka, corr = F)
ORand95CI(lowered_necka, confint_method = "wald")
## no effect. I also checked using the last 2 days of each phase. Same results

#PR vs baseline
lowered_neckb <- glmer(NECK_ABAIXO ~ GRUPO + (1|NOME), 
                    data = data_wth,
                    family = binomial)
summary(lowered_neckb, corr = F) #also a tendency, but see CI below
ORand95CI(lowered_neckb, confint_method = "profile")
```
Horses were less likely to show lowered neck during the exp phase compared to
baseline. However PR did not differ from control.


**TAIL MOVEMENT**

```{r}
#PR vs control
movin_taila <- glmer(RABO_MOVIMENTO ~ GRUPO + (1|NOME), 
                    data = data_btw,
                    family = binomial)
summary(movin_taila, corr = F)
## no effect. I also checked using the last 2 days of each phase. Same results

#PR vs baseline
movin_tailb <- glmer(RABO_MOVIMENTO ~ GRUPO + (1|NOME), 
                    data = data_wth,
                    family = binomial)
summary(movin_tailb, corr = F)
## no effect. I also checked using the last 2 days of each phase. Same results
```
**No effects on tail movement**

**RESTING**
This variable was categorized as "no rest", "part time resting" and "resting all
time". We will be looking at no rest and resting all time.

```{r}
#PR vs control
model_rest_alla <- glmer(as.factor(rest_alltime) ~ GRUPO + (1|NOME),
                    data = data_btw,
                    family = binomial)
summary(model_rest_alla, corr = F)
## model did not fit

#PR vs baseline
model_rest_allb <- glmer(as.factor(rest_alltime) ~ GRUPO + (1|NOME), 
                    data = data_wth,
                    family = binomial)
summary(model_rest_allb, corr = F)
## no effect. I also checked using the last 2 days of each phase. Same results

# NO REST
#PR vs control
model_no_rest_a <- glmer(as.factor(rest_never) ~ GRUPO + (1|NOME),
                    data = data_btw,
                    family = binomial)
summary(model_no_rest_a, corr = F)
## model did not fit

#PR vs baseline
model_no_rest_b <- glmer(as.factor(rest_never) ~ GRUPO + (1|NOME), 
                    data = data_wth,
                    family = binomial)
summary(model_no_rest_b, corr = F)
## no effect. I also checked using the last 2 days of each phase. Same results
```
**No effect on resting behavior**

### Facial expression

Facial action units were assessed using the ethogram by EquiFACS (Wathan et al
2015). 

#### Inter observer reliability

Because facial expressions are subjective we need to assess interobserver
reliability. There are several methodologies for that, but a simple one may be 
the use or ICC from the irr package that allows for checking reliability of
continuous variables. We will check if overall reliability of frequency and
duration os FACS from 5 randomly select 60 sec video clips.
```{r}
icc <-read_csv("ICC_laize_lais.csv")
event_freq <- select(icc, N_Laize, N_Lais)
duration <- select(icc, Duration_Laize, Duration_Lais) %>% 
  filter(!is.na(Duration_Laize))

#ICC
icc(event_freq, model="twoway", type="consistency")

```
There is consistency (ICC = 0.72) in the frequency of FACS observed by the two
raters.

### EquiFACS comparisons

60 sec video clips were analyzed for extracting all FACS expressed by the animals
in all phases of the study. Because the videos were collected after immediately
after control or positive reinforcement we may use the last two day of each
treatment. Most of the FACS variables will not be normally distributed, however
all of them are continuous and linear regressions are quite robust to deal with 
this type of data. Beside this analysis a second approach will also be
used to further investigate the effects of treatment: for count data, mixed poisson
will be used, while for continuous data that was extremely skewed we will drop 
all zero values and re-run the mixed linear regression or use a gamma regression
(as last resource).

```{r, warning=FALSE, message=FALSE}
df_facs <- FACS %>% filter(GROUP != "Baseline")


#AU101 - state event
summary(df_facs$AU101)
hist(df_facs$AU101) 
AU101 <- lmer(AU101 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU101, corr = F) #suppressing output
  emmeans(AU101, pairwise ~ GROUP)
  AU101_p <- glmer(AU101 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU101_p, corr = F)


#AU101 - point event 
  AU101_f <- lmer(AU101_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU101_f, corr = F) #supress output
  emmeans(AU101_f, pairwise ~ GROUP)

  AU101_fp <- glmer(AU101_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU101_fp, corr = F)
  ORand95CI(AU101_fp, confint_method = "wald")

#AU143 - point event
hist(df_facs$AU143)
  AU143 <- lmer(AU143 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU143, corr = F) #supress output
  emmeans(AU143, pairwise ~ GROUP)

  AU143_p <- glmer(AU143 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU143_p, corr = F)
  exp(fixef(AU143_p))

#AU145 - point event
hist(df_facs$AU145) #pretty normal, no need for poisson
  AU145 <- lmer(AU145 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU145, corr = F) #supressing output
  emmeans(AU145, pairwise ~ GROUP)
  

#AU47 - point event
hist(df_facs$AU47) #more normal than skewed
  AU47 <- lmer(AU47 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU47, corr = F) #supressing output
  emmeans(AU47, pairwise ~ GROUP)
  #AU47_p <- glmer(AU47 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  #summary(AU47_p, corr = F) #supressing output, results not diff than lmer

#AU5 - state event
hist(df_facs$AU5) #skewed
  AU5 <- lmer(AU5 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU5, corr = F)
  emmeans(AU5, pairwise ~ GROUP)
  #poisson model doesnt converge
  #AU5_p <- glmer(AU5 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  #summary(AU5_p, corr = F)
  #ORand95CI(AU5_p, confint_method = "wald")
  
#AU5 point event
  AU5_f <- lmer(AU5_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU5_f, corr = F) #supress output
  emmeans(AU5_f, pairwise ~ GROUP)
  
  AU5_fp <- glmer(AU5_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU5_fp, corr = F)
  ORand95CI(AU5_fp, confint_method = "wald")
  
  
#AD1 - state event
summary(df_facs$AD1)
  hist(df_facs$AD1) #skewed, must run poisson reg
  AD1 <- lmer(AD1 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AD1, corr = F) #
  emmeans(AD1, pairwise ~ GROUP)
  AD1_p <- glmer(AD1 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AD1_p, corr = F) 
  ORand95CI(AD1_p, confint_method = "wald")

#AD1 - point event
  AD1_f <- lmer(AD1_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AD1_f, corr = F) #supress output
  emmeans(AD1_f, pairwise ~ GROUP)
  
  AD1_fp <- glmer(AD1_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AD1_fp, corr = F)
  ORand95CI(AD1_fp, confint_method = "wald")
  
  
#AU10
summary(df_facs$AU10) #behaviour not observed in any horse

#AU12 - state event
hist(df_facs$AU12) #skewed, must check poission
  AU12 <- lmer(AU12 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU12, corr = F)
  emmeans(AU12, pairwise ~ GROUP) 
  AU12_p <- glmer(AU12 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU12_p, corr = F) 
  ORand95CI(AU12_p, confint_method = "wald")

#AU12 - point event
  AU12_f <- lmer(AU12_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU12_f, corr = F) #supress output
  emmeans(AU12_f, pairwise ~ GROUP)
  
  AU12_fp <- glmer(AU12_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU12_fp, corr = F)
  ORand95CI(AU12_fp, confint_method = "wald")


#AU113 - state event
summary(df_facs$AU113) #behavior not observed in any horse.

#AUH13 - state event
hist(df_facs$AUH13) #must check poisson
  AUH13 <- lmer(AUH13 ~ GROUP + (1|NAME), data = df_facs)
  summary(AUH13, corr = F)
  emmeans(AUH13, pairwise ~ GROUP)
  #AUH13_p <- glmer(AUH13 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  #summary(AUH13_p, corr = F) #doesnt converge
  #ORand95CI(AUH13_p, confint_method = "wald")

#AUH13 - point event
  AUH13_f <- lmer(AUH13_F ~ GROUP + (1|NAME), data = FACS3)
  #summary(AUH13_f, corr = F) #supress output
  emmeans(AUH13_f, pairwise ~ GROUP)
  
  AUH13_fp <- glmer(AUH13_F ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AUH13_fp, corr = F)
  ORand95CI(AUH13_fp, confint_method = "wald")
  
  
#AU16 - point event
hist(df_facs$AU16) #highly skewed
  AU16 <- lmer(AU16 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU16, corr = F)
  emmeans(AU16, pairwise ~ GROUP)
  AU16_p <- glmer(AU16 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU16_p, corr = F)

#AD160 (=AD160) - state event
hist(df_facs$AD160) #no need for poisson
  AU160 <- lmer(AD160 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU160, corr = F)
  emmeans(AU160, pairwise ~ GROUP) 

#AD160 (=AD160) - point event #Poisson model doesn't converge properly narrow CI
  AD160_f <- lmer(AD160_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AD160_f, corr = F) #supress output
  emmeans(AD160_f, pairwise ~ GROUP)
  
  #AD160_fp <- glmer(AD160_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  #summary(AD160_fp, corr = F)
  #ORand95CI(AD160_fp, confint_method = "wald")

#AU17 - point event
hist(df_facs$AU17) #highly skewed
  AU17 <- lmer(AU17 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU17, corr = F)
  emmeans(AU17, pairwise ~ GROUP)
  AU17_p <- glmer(AU17 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU17_p, corr = F)
  ORand95CI(AU17_p, confint_method = "wald")
  
#AU18 - state event
hist(df_facs$AU18) #must check poisson
AU18 <- lmer(AU18 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU18, corr = F)
  emmeans(AU18, pairwise ~ "GROUP")
  AU18_p <- glmer(AU18 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU18_p, corr = F)
  ORand95CI(AU18_p, confint_method = "wald")

#AU18 - point event
  AU18_f <- lmer(AU18_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU18_f, corr = F) #supress output
  emmeans(AU18_f, pairwise ~ GROUP)
  
  AU18_fp <- glmer(AU18_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU18_fp, corr = F)
  ORand95CI(AU18_fp, confint_method = "wald")
  
#AU122 - state event
hist(df_facs$AU122) #skewed must check poisson
  AU122 <- lmer(AU122 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU122, corr = F)
  emmeans(AU122, pairwise ~ GROUP)
  AU122_p <- glmer(AU122 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU122_p, corr = F)
  ORand95CI(AU122_p, confint_method = "wald")

#AU122 - point event
  AU122_f <- lmer(AU122_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU122_f, corr = F) #supress output
  emmeans(AU122_f, pairwise ~ GROUP)
  
  AU122_fp <- glmer(AU122_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU122_fp, corr = F)
  ORand95CI(AU122_fp, confint_method = "wald")
  
#AU24 - point event
hist(df_facs$AU24) #highly skewed
  AU24 <- lmer(AU24 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU24, corr = F)
  emmeans(AU24, pairwise ~ GROUP)
  AU24_p <- glmer(AU24 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU24_p, corr = F) #not signficant

#AU25 - state event
hist(df_facs$AU25) #skewed
  AU25 <- lmer(AU25 ~ GROUP + (1|NAME), data = df_facs)
  summary(AU25, corr = F)
  emmeans(AU25, pairwise ~ GROUP) 
  AU25_p <- glmer(AU25 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU25_p, corr = F)
  ORand95CI(AU25_p, confint_method = "wald")

#AU25 - point event
  AU25_f <- lmer(AU25_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU25_f, corr = F) #supress output
  emmeans(AU25_f, pairwise ~ GROUP)
  AU25_fp <- glmer(AU25_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU25_fp, corr = F)
  ORand95CI(AU25_fp, confint_method = "wald")
  
  
#AU26 - state event
hist(df_facs$AU26) #skewed must check poisson
  AU26 <- lmer(AU26 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AU26, corr = F) #suppress output
  emmeans(AU26, pairwise ~ GROUP) 
  AU26_p <- glmer(AU26 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AU26_p, corr = F)
  ORand95CI(AU26_p, confint_method = "wald")
  
#AU26 - point event
  AU26_f <- lmer(AU26_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU26_f, corr = F) #supress output
  emmeans(AU26_f, pairwise ~ GROUP)
  AU26_fp <- glmer(AU26_F ~ GROUP + (1|NAME), data = FACS2, family = "poisson")
  summary(AU26_fp, corr = F) #tendency
  ORand95CI(AU26_fp, confint_method = "wald") 
  
  
#AU27 - state event
hist(df_facs$AU27)
AU27 <- lmer(AU27 ~ GROUP + (1|NAME), data = df_facs)
summary(AU27, corr = F) #did not converge, control did not show this behavior
emmeans::emmeans(AU27, specs = "GROUP") 

#AU27 - point event
  AU27_f <- lmer(AU27_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(AU27_f, corr = F) #did not converge, control did not show this behavior
  emmeans(AU27_f, pairwise ~ GROUP)
  
#EAE101 (=AD101) - state event
hist(df_facs$AD101) #skewed
  AD101 <- lmer(AD101 ~ GROUP + (1|NAME), data = df_facs)
  summary(AD101, corr = F)
  emmeans(AD101, pairwise ~ GROUP)
  AD101_p <- glmer(AD101 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AD101_p, corr = F)
  ORand95CI(AD101_p, confint_method = "wald")

#EAE101 (=AD101) - state event
  EAE101_f <- lmer(AD101_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(EAE101_f, corr = F) #supress output
  emmeans(AU101_f, pairwise ~ GROUP)
  #model doesnt converge
  #EAE101_fp <- glmer(AD101_F ~ GROUP + (1|NAME), data = FACS2, 
   #                  family = "poisson")
  #summary(EAE101_fp, corr = F)
  #ORand95CI(EAE101_fp, confint_method = "wald")  
  
  
#EAE102 (=AD102) - state event
hist(df_facs$AD102) #highly skewed
  AD102 <- lmer(AD102 ~ GROUP + (1|NAME), data = df_facs)
  summary(AD102, corr = F)
  emmeans(AD102, pairwise ~ GROUP)
  AD102_p <- glmer(AD102 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AD102_p, corr = F)
  ORand95CI(AD102_p, confint_method = "wald")
  
#EAE102 (=AD102) - point event
  EAE102_f <- lmer(AD102_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(EAE102_f, corr = F) #supress output
  emmeans(EAE102_f, pairwise ~ GROUP)
  #model doesnot converge
  #EAE102_fp <- glmer(AD102_F ~ GROUP + (1|NAME), data = FACS2, 
   #                  family = "poisson")
  #summary(EAE102_fp, corr = F)
  #ORand95CI(EAE102_fp, confint_method = "wald")

#EAE103 (=AD103) - state event
hist(df_facs$AD103) 
AD103 <- lmer(AD103 ~ GROUP + (1|NAME), data = df_facs)
summary(AD103, corr = F)
emmeans(AD103, specs = "GROUP") #PR did not show this behavior
#EAE103 - Point event
EAE103_f <- lmer(AD103_F ~ GROUP + (1|NAME), data = FACS2)
  summary(EAE103_f, corr = F) #supress output
  emmeans(EAE103_f, pairwise ~ GROUP)

#EAE104 (=AD104) - State event
hist(df_facs$AD104) #highly skewed
  AD104 <- lmer(AD104 ~ GROUP + (1|NAME), data = df_facs)
  #summary(AD104, corr = F) #suppress output
  emmeans(AD104, pairwise~GROUP)
  AD104_p <- glmer(AD104 ~ GROUP + (1|NAME), data = df_facs, family = "poisson")
  summary(AD104_p, corr = F)
  ORand95CI(AD104_p, confint_method = "wald")
  
#EAE104 (=AD104) - point event
  EAE104_f <- lmer(AD104_F ~ GROUP + (1|NAME), data = FACS2)
  #summary(EAE104_f, corr = F) #supress output
  emmeans(EAE104_f, pairwise ~ GROUP)
  EAE104_fp <- glmer(AD104_F ~ GROUP + (1|NAME), data = FACS2, 
                    family = "poisson")
  summary(EAE104_fp, corr = F)
  ORand95CI(EAE104_fp, confint_method = "wald")
  
  #AD29 and AD30
#Both were not shown by any horse - no analysis needed
  
#AD133 - State event
  hist(FACS3$AD133_D) #highly skewed
  AD133 <- lmer(AD133_D ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD133, corr = F) #suppress output
  emmeans(AD133, pairwise~GROUP)
  AD133_p <- glmer(AD133_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AD133_p, corr = F)
  ORand95CI(AD133_p, confint_method = "wald")
  
#AD133 - point event
  hist(FACS3$AD133_F)
  AD133_f <- lmer(AD133_F ~ GROUP + (1|NAME), data = FACS3)
  summary(AD133_f, corr = F) #supress output
  emmeans(AD133_f, pairwise ~ GROUP)
  AD133_fp <- glmer(AD133_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD133_fp, corr = F)
  ORand95CI(AD133_fp, confint_method = "wald")
  
#AD38 - State event
  hist(FACS3$AD38_D) #highly skewed
  AD38 <- lmer(AD38_D ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD38, corr = F) #suppress output
  emmeans(AD38, pairwise~GROUP)
  AD38_p <- glmer(AD38_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AD38_p, corr = F)
  ORand95CI(AD38_p, confint_method = "wald")
  
#AD38 - point event
  hist(FACS3$AD38_F)
  AD38_f <- lmer(AD38_F ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD38_f, corr = F) #supress output
  emmeans(AD38_f, pairwise ~ GROUP)
  AD38_fp <- glmer(AD38_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD38_fp, corr = F)
  ORand95CI(AD38_fp, confint_method = "wald")
  
#AD55 - State event
  hist(FACS3$AD55_D) #highly skewed
  AD55 <- lmer(AD55_D ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD55, corr = F) #suppress output
  emmeans(AD55, pairwise~GROUP)
  AD55_p <- glmer(AD55_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AD55_p, corr = F)
  ORand95CI(AD55_p, confint_method = "wald")
  
#AD55 - point event
  hist(FACS3$AD55_F)
  AD55_f <- lmer(AD55_F ~ GROUP + (1|NAME), data = FACS3)
  summary(AD55_f, corr = F) #supress output
  emmeans(AD55_f, pairwise ~ GROUP)
  AD55_fp <- glmer(AD55_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD55_fp, corr = F)
  ORand95CI(AD55_fp, confint_method = "wald")
  
  
#AD56 - State event
  hist(FACS3$AD56_D) #highly skewed
  AD56 <- lmer(AD56_D ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD56, corr = F) #suppress output
  emmeans(AD56, pairwise~GROUP)
  AD56_p <- glmer(AD56_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AD56_p, corr = F)
  ORand95CI(AD56_p, confint_method = "wald")
  
#AD56 - point event
  hist(FACS3$AD56_F)
  AD56_f <- lmer(AD56_F ~ GROUP + (1|NAME), data = FACS3)
  summary(AD56_f, corr = F) #supress output
  emmeans(AD56_f, pairwise ~ GROUP)
  AD56_fp <- glmer(AD56_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD56_fp, corr = F)
  ORand95CI(AD56_fp, confint_method = "wald")
  
  
  #AD57 - State event
  hist(FACS3$AD57_D) #highly skewed
  AD57 <- lmer(AD57_D ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD57, corr = F) #suppress output
  emmeans(AD57, pairwise~GROUP)
  #model failed to converger
  #AD57_p <- glmer(AD57_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  #summary(AD57_p, corr = F)
  #ORand95CI(AD57_p, confint_method = "wald")
  
#AD57 - point event
  hist(FACS3$AD57_F)
  AD57_f <- lmer(AD57_F ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD57_f, corr = F) #supress output
  emmeans(AD57_f, pairwise ~ GROUP)
  AD57_fp <- glmer(AD57_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD57_fp, corr = F)
  ORand95CI(AD57_fp, confint_method = "wald")
  
#AD58 - State event
  hist(FACS3$AD58_D) #highly skewed
  AD58 <- lmer(AD58_D ~ GROUP + (1|NAME), data = FACS3)
  summary(AD58, corr = F) #suppress output
  emmeans(AD58, pairwise~GROUP)
  AD58_p <- glmer(AD58_D ~ GROUP + (1|NAME), data = FACS3, family = "poisson")
  summary(AD58_p, corr = F)
  ORand95CI(AD58_p, confint_method = "wald")
  
#AD58 - point event
  hist(FACS3$AD58_F)
  AD58_f <- lmer(AD58_F ~ GROUP + (1|NAME), data = FACS3)
  #summary(AD58_f, corr = F) #supress output
  emmeans(AD58_f, pairwise ~ GROUP)
  AD58_fp <- glmer(AD58_F ~ GROUP + (1|NAME), data = FACS3, 
                    family = "poisson")
  summary(AD58_fp, corr = F)
  ORand95CI(AD58_fp, confint_method = "wald")
```

**Several FACS differed between phases.** Note that some models did not converge
and model fit was not ideal in many instances as the distribution was wacky...
Other models could have been used, eg, gamma truncated though to my knowledge we
do not have this implemented for mixed models in R.

## END

