---
title: "COVID-19-related research data availability and quality according to the FAIR principles: Analyses"
author: "Ahmad Sofi-Mahmudi"
format: 
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
        #html:
        #        toc: true
        #        toc-expand: false
        #        html-math-method: katex
editor: visual
---

# Aim

We aimed to determine the availability of open data in dental research and to assess compliance with the FAIR principles (or FAIRness) of shared COVID-19-related research data.

# Results

First, loading the needed packages:

```{r}
pacman::p_load(tidyverse, 
               here, # for REPRODUCIBILITY
               janitor, # for data cleaning, tables and etc
               scales, 
               patchwork, 
               knitr,
               gtsummary,
               forcats,
               tidyr,
               treemap,
               DiagrammeR)
```

```{r}
here()
```

And then, loading the datasets:

```{r}
fairness = read.csv(here("data", "covid_fairness_results_final.csv"))
```

## **General characteristics**

Let's take a look at the number of papers published in each year and overall:

```{r}
fairness %>% tbl_summary(include = c(pubYear),
                         digits = list(all_categorical() ~ c(0, 1)))
```

What about article types?

```{r}
fairness %>% tbl_summary(
  include = c(is_reseach, is_review, is_observational, is_rct, is_letter),
  digits = list(all_categorical() ~ c(0, 1))
)
```

Let's take a look at the top journals in terms of the number of papers in our dataset:

```{r}
fairness %>% 
  mutate(journalTitle = fct_lump(journalTitle, n = 5)) %>%
  tbl_summary(
    include = journalTitle,
    digits = list(all_categorical() ~ c(0, 1)),
    sort = all_categorical() ~ "frequency"
  )
```

## **FAIRness results**

The number and proportion of papers with score of 1:

```{r}
fairness %>% 
  mutate(FAIR = as.factor(FAIR)) %>%
        tbl_summary(include = c(FAIR), 
                         digits = list(all_categorical() ~ c(0, 1)))
```

We remove the repositories with a FAIR score of 1 since they were either inaccessible or had nothing in them.

```{r}
fairness <- fairness %>% 
  filter(FAIR != 1)
```

Now, let's take a look at mean and SD for each component and subcomponent of FAIRness:

```{r}

# dat <-  data.frame()
# row  <-  15
# 
# while(row < 32){
#         mean1 = round(mean(select(fairness, row)[, 1], na.rm = T), 1)
#         sd1 = round(sd(select(fairness, row)[, 1], na.rm = T), 2)
#         range1 = range(select(fairness, row))[1]
#         range2 = range(select(fairness, row))[2]
#         over1 = paste0(mean1, " (", sd1, ")")
#         over2 = paste0(range1, "–", range2)
#         dat[row-14, 1:3] = c(names(fairness)[row], over1, over2)
#         row = row+1
# }
# 
# dat = dat[c(8, 3:7, 1:2, 9:17), ]
# names(dat) = c("Metric", "Mean (SD)", "Range")
# 
# kable(dat, row.names = F)
```

We now add the FAIR levels:

-   0: incomplete,

-   1: initial,

-   \>1: moderate,

-   max: advanced

```{r}
fairness <-  fairness %>% mutate(Findability = ifelse(`F` == 0, "incomplete",
                                                ifelse(`F` == 1, "initial",
                                                       ifelse(`F` == 7, "advanced", "moderate"))), 
                               Accessibility = ifelse(`A` == 0, "incomplete",
                                                ifelse(`A` == 1, "initial",
                                                       ifelse(`A` == 3, "advanced", "moderate"))),
                               Interoperability = ifelse(`I` == 0, "incomplete",
                                                ifelse(`I` == 1, "initial",
                                                       ifelse(`I` == 4, "advanced", "moderate"))),
                               Reusability = ifelse(`R` == 0, "incomplete",
                                                ifelse(`R` == 1, "initial",
                                                       ifelse(`R` == 8, "advanced", "moderate"))))
```

And now, a summary of the levels:

```{r}
fairness %>% tbl_summary(
  include = c(Findability, Accessibility, Interoperability, Reusability),
  digits = list(all_categorical() ~ c(0, 1)),
  sort = all_categorical() ~ "frequency"
)
```

### **FAIRness by article type**

First, creating a new variable for article types:

```{r}
fairness = fairness %>% mutate(
        article_type = case_when(
                is_rct ~ "rct",
                is_observational ~ "observational",
                is_review ~ "review",
                is_letter ~ "letters",
                TRUE ~ "other_research"
        ))
```

And then, mean and SD of FAIR score by article type:

```{r}
fairness %>% select(article_type, FAIR) %>%
        tbl_summary(by = article_type,
                    statistic = all_continuous() ~ "{mean} ({sd})",
                    digits = all_continuous() ~ 2) %>% add_p()
```

### **FAIRness by journal area**

First, creating a new variable for journal area:

```{r}
fairness <-  fairness %>% mutate(
        journal_area = case_when(
                is_biochem ~ "biochemistry",
                is_dentistry ~ "dentistry",
                is_immunology ~ "immunology",
                is_neuroscience ~ "neuroscience",
                is_nursing ~ "nursing",
                is_pharmacology ~ "pharmacology",
                is_psychology ~ "psychology",
                is_medicine ~ "medicine",
                is_health ~ "health",
                is_multidisciplinary ~ "multidisciplinary",
                TRUE ~ "other"
        ))
```

And then, mean and SD of FAIR score by article type:

```{r}
fairness %>% select(journal_area, FAIR) %>%
        tbl_summary(by = journal_area,
                    statistic = all_continuous() ~ "{mean} ({sd})",
                    digits = all_continuous() ~ 2) %>% add_p()
```

### **FAIRness by repository**

First, creating a new variable for repository:

```{r}
fairness <-  fairness %>% mutate(
        repository = case_when(
                grepl("figshare|6084", data_url) ~ "figshare",
                grepl("datadryad|5061", data_url) ~ "dryad",
                grepl("zenodo|5281", data_url) ~ "zenodo",
                grepl("dataverse|7910", data_url) ~ "dataverse",
                grepl("dataversenl|34894", data_url) ~ "dataversenl",
                grepl("osf|17605", data_url) ~ "osf",
                grepl("mendeley|17632", data_url) ~ "mendeley",
                grepl("gigadb|5524", data_url) ~ "gigadb",
                grepl("openneuro|18112", data_url) ~ "openneuro",
                grepl("github", data_url) ~ "github"
        )
)
```

And then, mean and SD of FAIR score by article type:

```{r}
fairness %>% select(repository, FAIR) %>%
        tbl_summary(by = repository,
                    statistic = all_continuous() ~ "{mean} ({sd})",
                    digits = all_continuous() ~ 2) %>% add_p()
```

### **The most influential factor**

The linear regression model for article_type, adjusted for sjr2020 and citedByCount:

```{r}
lm_article = lm(FAIR~article_type+sjr2020+citedByCount, data = fairness)

summary(lm_article)
```

R2 = 0.001518

```{r}
lm_journal = lm(FAIR~journal_area+sjr2020+citedByCount, data = fairness)

summary(lm_journal)
```

R2=0.01614

```{r}
lm_repository = lm(FAIR ~ repository + sjr2020 + citedByCount, data = fairness)

summary(lm_repository)
```

R2=0.8088

```{r}
lm_all = lm(FAIR ~ article_type + journal_area + repository + sjr2020 + citedByCount,
            data = fairness)

summary(lm_all)
```

R2=0.8116

```{r}
lm_all <- gtsummary::tbl_regression(lm_all)
lm_repository <- gtsummary::tbl_regression(lm_repository)
lm_journal <- gtsummary::tbl_regression(lm_journal)
lm_article <- gtsummary::tbl_regression(lm_article)
```

```{r}

 tbl_merge(
    tbls = list(lm_repository, 
                lm_journal, 
                lm_article, 
                lm_all),
    tab_spanner = c("**Repository**", 
                    "**Journal**", 
                    "**Article**", 
                    "**All**")
  )
 

```

## Figures

### Figure 1: Treemap

```{r}

treemap <-  data.frame(oa = c(paste("Non-open-access", "87,984", sep = '\n'), "Open-access", "Open-access", "Open-access"),
                     shared = c("All", paste("Not shared", "236,475", sep = '\n'), "Shared", "Shared"),
                     repo = c(NA, NA, paste("General repository", "8,015", sep = '\n'), paste("Other", "12,858", sep = '\n')),
                     Papers = c(87984, 236475, 8015, 12858))

fig1  <-  treemap(treemap,
             index=c("oa","shared","repo"),
             vSize="Papers",
             type="index",
             palette = "Set2",
             align.labels=list(
                 c("center", "center"), 
                 c("right", "bottom")
             )  
) 

#tiff("figures/Figure 1.tiff", width = 6, height = 4, units = "in", res = 300, compression = "lzw")
#fig1
#dev.off()
```

### Figure 2: Flow diagram

```{r}
fig2 <-  grViz(diagram = "digraph flowchart {
  node [fontname = arial, shape = box, color = grey, style = filled]
  tab1 [label = '@@1']
  tab2 [label = '@@2']
  tab3 [label = '@@3']
  tab4 [label = '@@4']
  tab5 [label = '@@5']
  tab6 [label = '@@6']
  tab7 [label = '@@7']
  
  tab1 -> tab2
  tab2 -> tab3
  tab2 -> tab4
  tab2 -> tab5
  tab2 -> tab6
  tab2 -> tab7;
}
  
  [1]: 'All open-access COVID-19-related articles in Europe PMC from 2020-01-01 to 2023-04-15 (N=257,348)'
  [2]: 'Articles shared their data in a general repository (n=6,180)'
  [3]: 'Research articles (n=5,580)'
  [4]: 'Reviews (n=182)'
  [5]: 'Observational studies (n=121)'
  [6]: 'RCTs (n=107)'
  [7]: 'Letters (n=64)'


  ")

#tiff("figures/Figure 2.tiff", width = 6, height = 2, units = "in", res = 300, compression = "lzw")
fig2
#dev.off()
```

### Figure 3: Compliance level

```{r}
data_long <-  gather(fairness, fairness, level, Findability:Reusability, factor_key=TRUE)

table_stack = as.data.frame(table(data_long$fairness, data_long$level))
table_stack$Percent = table_stack$Freq/5700*100
names(table_stack) = c("FAIR Metric", "FAIR Level", "Frequency", "Percent")
table_stack$`FAIR Level` = factor(table_stack$`FAIR Level`, levels = c("advanced", "moderate", "initial", "incomplete"))

fig3 = table_stack %>% ggplot() + 
        aes(x = `FAIR Metric`, y = Percent, fill = `FAIR Level`) +
        geom_bar(position='stack', stat='identity') +
        scale_y_continuous(labels = scales::percent_format(scale = 1)) +
        scale_fill_manual(values=c('green', 'lightgreen', 'yellow', 'red')) +
        theme_minimal()

#tiff("figures/Figure 3.tiff", width = 6, height = 4, units = "in", res = 300, compression = "lzw")
fig3
#dev.off()
```

### Figure 4: Monthly trend

First, we add the real publication year and month to the datasets. The real publication year/month is the year/month the paper was first appear online stored in firstPublicationDate column.

```{r}
fairness$YearMonth <- format(as.Date(fairness$firstPublicationDate), "%Y-%m")
```

And now, creating a trend dataset:

```{r}
monthly_trend = 
        fairness %>% 
        group_by(YearMonth) %>%
        summarize_at(vars("FAIR", "F", "A", "I", "R"), mean)

monthly_trend = gather(monthly_trend, indicator, value, FAIR:R, factor_key = T)
```

Now, we can draw the trend:

```{r}
fig4 =
        monthly_trend %>% 
        ggplot() +
        aes(x = YearMonth, 
            y = value,
            group = indicator,
            color = indicator) +
        geom_line(linewidth = 0.75) +
        scale_y_continuous(breaks = seq(0, 15, by = 1),
                           labels = c("", 1, "", "", "", 5, "", "", "", "", 10, "", "", "", 14, 15))+
        scale_color_discrete(name = NULL) +
        labs(y = "Score", 
             x = "Month") +
        theme(panel.grid.minor = element_blank(),
              axis.text.x = element_text(angle = 55, vjust = 0.9, hjust=1)
              )

#tiff("figures/Figure 4.tiff", width = 8, height = 4, units = "in", res = 300, compression = "lzw")
fig4
#dev.off()
```

### Figure 4: second version

```{r}
# Show the year only when month = Jan
label_year_if_jan <- function(date) {
  ifelse(format(date, "%m") == "01", format(date, "%b\n%Y"), format(date, "%b"))
}

```

```{r}

fair_monthly <- monthly_trend %>%
  filter(indicator != "FAIR") %>%
  mutate(YearMonth = as.Date(paste0(YearMonth, "-01"))) %>%
  ggplot(aes(x = YearMonth, y = value, group = indicator, color = indicator)) +
    geom_line(linewidth = 0.75) +
    scale_y_continuous(breaks = seq(0, 15, by = 1),
                       labels = c("", 1, "", "", "", 5, "", "", "", "", 10, "", "", "", 14, 15)) +
    scale_x_date(date_breaks = "1 month", labels = label_year_if_jan) +
    scale_colour_viridis_d() +
    labs(y = "Score", x = "", color = "Indicator") +
    theme(panel.grid.minor = element_blank(),
          axis.text.x = element_blank()) + 
    theme_minimal() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  facet_grid(. ~ indicator)
```

```{r}

fair_year <- monthly_trend %>%
  filter(indicator == "FAIR") %>%
  mutate(YearMonth = as.Date(paste0(YearMonth, "-01"))) %>%
  ggplot(aes(x = YearMonth, y = value, group = indicator, color = indicator)) +
    geom_line(linewidth = 0.75) +
    scale_y_continuous(breaks = seq(0, 15, by = 1),
                       labels = c("", 1, "", "", "", 5, "", "", "", "", 10, "", "", "", 14, 15)) +
    scale_x_date(date_breaks = "1 month", labels = label_year_if_jan) +
    scale_colour_viridis_d(option = "magma", begin = .5) +
    labs(y = "Score", x = "Month", color = "") +
    theme(panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 55, vjust = 0.9, hjust=1)) + 
    theme_minimal()
```

```{r}
fair_monthly / fair_year
```
